resistanceModelNotesAndy

### 2016-01-15 effectOfIndividualInputs.Rmd
plots of effect of individual inputs

---
title: "sensiAnPaper1All"
output: html_document
---

### All code for the sensitivity analysis for the first paper

The main question is under which scenarios are mixtures better than sequential ?

This breaks down into these subsidiary quetsions :
1. Under which scenarios is the time to resistance for either insecticide quicker for

```{r}

  #library(devtools)
  #install_github('AndySouth/resistance')
  library(resistance)

  outFolder <- "C:\\Dropbox\\resistanceResults\\"
```

<!-- eval=FALSE to stop slow code being executed --> 
```{r, eval=FALSE}

  #default setting of the random seed in sensiAnPaperPart() should make the input files mostly the same with exception of exposure

  #nScenarios <- 100
  nScenarios <- 1000
  ## do model runs
  listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture' )
  listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1' )
  listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2' )

  ## save results objects as rda for re-use
  save(listOutMix, file=paste0(outFolder,'listOutMix.rda'))
  save(listOutI1, file=paste0(outFolder,'listOutI1.rda'))
  save(listOutI2, file=paste0(outFolder,'listOutI2.rda'))

```{r}
  ## to load previously saved runs, e.g. these that I added 100 onto filename
  load(file=paste0(outFolder,'listOutMix_1000.rda'))
  load(file=paste0(outFolder,'listOutI1_1000.rda'))
  load(file=paste0(outFolder,'listOutI2_1000.rda'))

### now want to calculate the times to reach critical points and add them on to the input file 
### for different insecticide strategies, sequential, mix1 and mix2
```{r}
  #1) sequential : time to resistance for each insecticide in isolation
  #inputs : inAndOutI1, inAndOutI2
  #find time to criticalPoint for insecticide1
  #find time to criticalPoint for insecticide2
  #add together
  resistPointsI1 <- findResistancePoints(listOutI1, locus=1)
  resistPointsI2 <- findResistancePoints(listOutI2, locus=2)

@@

### calculate the difference between sequential & mixture scenarios
```{r}

  require('plyr')

  ## is mix1 better than sequential ?
  #T/F
  resistBetterMix1SeqBoolean <- resistPointsMix1 > resistPointsSeq
  #convert to 0/1
  resistBetterMix1Seq <- plyr::aaply(resistBetterMix1SeqBoolean,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix1Seq) <- gsub('gen','betterMix1Seq', rownames(resistBetterMix1Seq))

  rowSums(resistBetterMix1Seq)/ncol(resistBetterMix1Seq)
  #100runs
  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
  #                .25                   .26                   .30   
  #1000runs
  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
  #               .178                  .232                  .238 
  #new100
  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
  #              0.18                 0.24                 0.25 
  #new1000
  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
  #             0.223                0.245                0.243

  ## is mix2 better than sequential ?
  #T/F
  resistBetterMix2SeqBoolean <- resistPointsMix2 > resistPointsSeq
  #convert to 0/1
  resistBetterMix2Seq <- plyr::aaply(resistBetterMix2SeqBoolean,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix2Seq) <- gsub('gen','betterMix2Seq', rownames(resistBetterMix2Seq))

  rowSums(resistBetterMix2Seq)/ncol(resistBetterMix2Seq)  
  #1000runs
  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
  #             0.471                0.568                0.551 
  #new100
  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
  #              0.58                 0.57                 0.52 
  #new1000
  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
  #             0.558                0.563                0.534

  ## is mix3 better than sequential ?
  #T/F
  resistBetterMix3SeqBoolean <- resistPointsMix3 > resistPointsSeq
  #convert to 0/1
  resistBetterMix3Seq <- plyr::aaply(resistBetterMix3SeqBoolean,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix3Seq) <- gsub('gen','betterMix3Seq', rownames(resistBetterMix3Seq))

  #tally runs where mix3 better than seq
  rowSums(resistBetterMix3Seq)/ncol(resistBetterMix3Seq)
  #100runs
  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
  #                  .83                   .75                   .72 
  #1000runs
  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
  #                .794                  .726                  .673  
  #new100
  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
  #              0.83                 0.79                 0.72 
  #new1000
  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
  #             0.820                0.747                0.677

  #check1
  #plot the 2 individual results & the mixture, just to check what is going on
  #x11()
#   plotallele.freq.andy(listOutMix)
#   #x11()
#   plotallele.freq.andy(listOutI1)
#   #x11()
#   plotallele.freq.andy(listOutI2)

  #I should be able to compare single scenarios
  #yes this works, but not too useful
  # plotallele.freq.andy( list(results=list(listOutMix$results[[1]], listOutI1$results[[1]] )) )
  #this is a modifdeied version of Beths function
  #allowing curtis fig2 to be applied to any scenario
  #plotcurtis_f2_generic( listOutMix$results[[1]], listOutI2$results[[1]], listOutI1$results[[1]] )

  #check2
#   #what is the absolute difference between mix3 & sequential ?
#   resistDifferenceMix3Seq <- resistPointsMix3 - resistPointsSeq
#   
#   #check3 is their a difference between the mix & the individual strategies
#   #looks mostly plausible, occasionally there are 0s
#   resistDifferenceMix1I1 <- resistPointsMix1 - resistPointsI1
#   resistDifferenceMix1I2 <- resistPointsMix1 - resistPointsI2

  ## is mix2 better than sequential ?  
  # I might expect it to be similar to mix3

  # comparing sensitivity inputs to those used by Beth to reproduce curtis fig2
#   inputCurtis <- createInputMatrix(FALSE)
#   inputSensiMix <- listOutMix$input
#   
#   input <- cbind(inputCurtis, inputSensiMix)
#   write.csv(input, file=paste0(outFolder,'inputCurtisSensiCompare.csv'))

### finding when mixtures are >20% better
### repeating calculation of the difference between sequential & mixture scenarios
```{r}

  require('plyr')

  ## is mix1 better than sequential ?
  #T/F
  resistBetterMix1SeqBoolean20 <- resistPointsMix1 > (1.2 * resistPointsSeq)
  #convert to 0/1
  resistBetterMix1Seq20 <- plyr::aaply(resistBetterMix1SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix1Seq20) <- gsub('gen','betterMix1Seq20', rownames(resistBetterMix1Seq20))

  rowSums(resistBetterMix1Seq20)/ncol(resistBetterMix1Seq20)
  #1000runs
 #betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
 #                 0.136                  0.175                  0.173 
  #new1000
 #betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
#                 0.163                  0.178                  0.182

  ## is mix2 better than sequential ?
  #T/F
  resistBetterMix2SeqBoolean20 <- resistPointsMix2 > (1.2 * resistPointsSeq)
  #convert to 0/1
  resistBetterMix2Seq20 <- plyr::aaply(resistBetterMix2SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix2Seq20) <- gsub('gen','betterMix2Seq20', rownames(resistBetterMix2Seq20))

  rowSums(resistBetterMix2Seq20)/ncol(resistBetterMix2Seq20)  
  #1000runs
#  betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
#                  0.241                  0.326                  0.313 
  #new1000
# betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
#                 0.283                  0.302                  0.283

  ## is mix3 better than sequential ?
  #T/F
  resistBetterMix3SeqBoolean20 <- resistPointsMix3 > (1.2 * resistPointsSeq)
  #convert to 0/1
  resistBetterMix3Seq20 <- plyr::aaply(resistBetterMix3SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix3Seq20) <- gsub('gen','betterMix3Seq20', rownames(resistBetterMix3Seq20))

  #tally runs where mix3 better than seq
  rowSums(resistBetterMix3Seq20)/ncol(resistBetterMix3Seq20)
  #1000runs
#  betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
#                  0.672                  0.560                  0.488   
  #new1000
# betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
#                 0.622                  0.536                  0.458

### bind results onto input file
```{r}

@@

  #get the inputs here to use in ggplot investigation of model responses to inputs
  #used in a later chunk
  ggInput <- treeInput

  #replace 0s & 1s with mixture/sequence to make plot clearer
  repMix <- function(x){
    x[x==0] <- 'sequence'
    x[x==1] <- 'mixture'
    x
  }

  resistBetterMix1Seq <- repMix(resistBetterMix1Seq)
  resistBetterMix2Seq <- repMix(resistBetterMix2Seq)
  resistBetterMix3Seq <- repMix(resistBetterMix3Seq)

  resistBetterMix1Seq20 <- repMix(resistBetterMix1Seq20)
  resistBetterMix2Seq20 <- repMix(resistBetterMix2Seq20)
  resistBetterMix3Seq20 <- repMix(resistBetterMix3Seq20)

  #bind results onto predictors
  treeInput <- rbind( treeInput, resistBetterMix1Seq )
  treeInput <- rbind( treeInput, resistBetterMix2Seq )
  treeInput <- rbind( treeInput, resistBetterMix3Seq )
  #add >20% results
  treeInput <- rbind( treeInput, resistBetterMix1Seq20 )
  treeInput <- rbind( treeInput, resistBetterMix2Seq20 )
  treeInput <- rbind( treeInput, resistBetterMix3Seq20 )

### do tree analysis
```{r, eval=TRUE}
  require(rpart.plot)

  #transpose
  treeInput <- t(treeInput)

  #convert to a dataframe
  treeInput <- data.frame(treeInput, stringsAsFactors = FALSE)

  #convert predictor columns back to numeric
  #following this above : replace 0s & 1s with mixture/sequence to make plot clearer
  filter <- !substr(names(treeInput),1,6) %in% 'better'
  treeInput[,filter] <- lapply(treeInput[,filter],as.numeric) 
  filter <- !filter
  treeInput[,filter] <- lapply(treeInput[,filter],as.factor)

  #create string with predictor names in
  treePredictorString <- paste(treePredictors, collapse="+")

  treeResponses <- c( rownames(resistBetterMix1Seq),
                      rownames(resistBetterMix2Seq),
                      rownames(resistBetterMix3Seq),
                      rownames(resistBetterMix1Seq20),
                      rownames(resistBetterMix2Seq20),
                      rownames(resistBetterMix3Seq20)
                      )

  #to do trees & plots for all response variables
  for( treeResponse in treeResponses )
  {
    x11()
    #png(paste0(outFolder,"tree",treeResponse,".png"))

    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
    #treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])

    #http://stackoverflow.com/questions/29197213/what-is-the-difference-between-rel-error-and-x-error-in-a-rpart-decision-tree
    #A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.

    cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]

    treePruned <- prune(tree, cp=cpOptimal)

    #to set box colours
    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")

    #adds in no. scenarios correctly and incorrectly classified
    #prp(tree, extra=1, varlen = 10, main=treeResponse)
    #prp(treePruned, extra=1, varlen = 10, main=treeResponse, col=cols, border.col=cols)
    #prp(treePruned, extra=1, varlen = 0, main=treeResponse, col=cols, border.col=cols)
    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)

    #dev.off()
    #savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')
    savePlot(paste0(outFolder,"tree",treeResponse,".jpg"), type='jpg')

    #trying to limit num levels a different way using maxdepth
    #tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class', maxdepth=3) 
    #adds in no. scenarios correctly and incorrectly classified
    #prp(tree, extra=1, varlen = 10, main=treeResponse)
    #savePlot(paste0(outFolder,"tree",treeResponse,"depth3.jpg"), type='jpg')

  }
```

### testing tree analysis for paper
```{r, eval=TRUE}

    x11()

    treeResponse <- "betterMix2Seq_cP0.5"

    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])

    #to set box colours
    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")

    #adds in no. scenarios correctly and incorrectly classified
    #prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)

    cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]

    treePruned <- prune(tree, cp=cpOptimal)

    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)

    #comparing it with 0/1 gives the same big tree
    treeInput$betterMix2Seq_cP0.5numeric <- as.numeric(treeInput$betterMix2Seq_cP0.5)

    treeResponse <- "betterMix2Seq_cP0.5numeric"

    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])

    #to set box colours
    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")

    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)

    #testing alternative pruning
    #A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.
    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])

```    
#   treeResponse <- 'betterMix1Seq_cP0.5'
#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
#   
#   #if get error number of rows of matrices must match (see arg 2)
#   #is because all vals are 0
# 
#   plot(tree) #plots tree
#   text(tree) #labels tree
#   #branch predictors 2 phis & P_2  
#     
#   #for mix3 (resistance to both)
#   x11()
#   treeResponse <- 'betterMix3Seq_cP0.5'
#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class')  
#   plot(tree) #plots tree
#   text(tree) #labels tree  
#   
#   #branch predictors 2 phis & P_1
#   
#   #or from rpart.plot
#   #prp(tree)
#   
#   #adds in no. scenarios correctly and incorrectly classified
#   prp(tree, extra=1, varlen = 10, main=treeResponse)
#   savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')

  #prp(tree, extra=1,box.col=c("pink","palegreen3")[tree$frame$yval],cex=0.5,yesno.yshift=-2)

    #susanas code
    #pruning tree
    #   treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
    #   plot(treePruned) #plots tree
    #   text(treePruned) #labels tree 
    # prp(prunned_tree, extra=1,box.col=c("pink", "palegreen3")[prunned_tree$frame$yval],cex=0.5,yesno.yshift=-2)
```

### investigating response of time-to-resistance to individual input parameters
### (while they all are varying so it may not work well)
```{r, eval=TRUE}

  #uses ggInput calculated in earlier trunk

  #first needs to transpose rows to cols
  ggInput_T <- t(ggInput)

@@

  #this one shows no discernible pattern
  ggplot(ggInsOuts, aes(x=effectiveness_insecticide1, y=gen_cP0.5, color=strategy)) + geom_point() + facet_wrap( ~ strategy) + geom_smooth()

@@

  for(i in names_inputs)
  {
    x11()

    y <- 'gen_cP0.5'
    #color <- 'strategy'
    #coloring by exposure shows its effect even when other params are varying
    color <- 'exposure'

    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
             geom_point() + 
             facet_wrap( ~ strategy) + 
             geom_smooth(aes_string(x=i, y=y, color=NULL)) )
  }

>>>>

---
  title: "effectOfIndividualInputs"
  output: pdf_document
  author: "Andy South"
  date: "`r Sys.Date()`"
---

# Investigation into effects of individual inputs on time to reach resistance thresholds.
# version 2 - in progress

## Based on 1000 runs with all inputs varying at once which explains the variability at particular input values.
## Within each figure the plots are divided into 6 sub-plots according to the insecticide use strategy and when the resistance threshold is reached
1. insecticide 1 : sole use
1. insecticide 2 : sole use
1. Mixture 1 : threshold reached for either insecticide in mixture
1. Mixture 2 : once threshold reached for either insecticide in mixture, switch to sole use of other until it too reaches threshold
1. Mixture 3 : threshold reached for both insecticides in mixture
1. Sequential : sole use of one insecticide, switch to other when threshold reached until it too reaches threshold

These plots show gen_cP0.2, which is the number of generations to reach 20% resistance.

Red dashed lines are a smoothed mean.

The plots show a set of points at 1000 generations. The model is run for 500 generations, any runs which have not reached the resistance threshold by this time are given a value of 1000. This has little effect on questions of whether mixtures or sequential use is better, but does effect the plots presented here.

```{r,echo=FALSE,results='hide'}

  library(resistance)

  outFolder <- "C:\\Dropbox\\resistanceResults\\"

  ## to load previously saved runs
  load(file=paste0(outFolder,'listOutMix_1000.rda'))
  load(file=paste0(outFolder,'listOutI1_1000.rda'))
  load(file=paste0(outFolder,'listOutI2_1000.rda'))
```

```{r,echo=FALSE,results='hide'}
### chunk copied from sensiAnPaper1All
### now want to calculate the times to reach critical points and add them on to the input file 
### for different insecticide strategies, sequential, mix1 and mix2

  #1) sequential : time to resistance for each insecticide in isolation
  #inputs : inAndOutI1, inAndOutI2
  #find time to criticalPoint for insecticide1
  #find time to criticalPoint for insecticide2
  #add together
  resistPointsI1 <- findResistancePoints(listOutI1, locus=1)
  resistPointsI2 <- findResistancePoints(listOutI2, locus=2)

@@

```{r,echo=FALSE,results='hide'}
### chunk copied from part of one in sensiAnPaper1All
### bind results onto input file

@@

  #get the inputs here to use in ggplot investigation of model responses to inputs
  #used in a later chunk
  ggInput <- treeInput
```

## This plot shows exposure which has the most noticeable pattern.
```{r, eval=TRUE, echo=FALSE, warning=FALSE}

  library(ggplot2)

  #uses ggInput calculated in earlier chunk

  #first needs to transpose rows to cols
  #ggInput_T <- t(ggInput)

@@

  #exposure has the most noticeable effect
  ggplot(ggInsOuts, aes(x=exposure, y=gen_cP0.25, color=strategy)) + 
    geom_point() + 
    facet_wrap( ~ strategy) + 
    #geom_smooth(aes(x=exposure, y=gen_cP0.5, color=NULL))
    geom_smooth(colour='red', linetype='dashed',size=0.5)

  #effectiveness of insecticide1 little noticeable effect
#   ggplot(ggInsOuts, aes(x=effectiveness_insecticide1, y=gen_cP0.25, color=strategy)) + 
#     geom_point() + 
#     facet_wrap( ~ strategy) + 
#     #geom_smooth(aes(x=effectiveness_insecticide1, y=gen_cP0.5, color=NULL)) 
#     geom_smooth(colour='red', linetype='dashed',size=0.5)

## In the following plots all inputs are plotted individually, the points are coloured by exposure, which allows us to see that, even as other inputs change, low times to resistance are associated with high exposure.
```{r, eval=TRUE, echo=FALSE}

@@

  for(i in names_inputs)
  {
    #x11()

    y <- 'gen_cP0.25'    
    #y <- 'gen_cP0.5'
    #color <- 'strategy'
    #coloring by exposure shows its effect even when other params are varying
    color <- 'exposure'

    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
             geom_point() + 
             facet_wrap( ~ strategy) + 
             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
             geom_smooth(colour='red', linetype='dashed',size=0.5) )
  }
